Profit Optimisation in Heterogeneous 2-link Parallel Games

### 2024-09-16 thesis/article.tex
Thesis: Adds Preliminaries, Network Games, Heterogeneous players
Started Heterogeneous players, WIP

\clearpage
\thispagestyle{empty}

@@

\chapter{Related Work}
TODO

@@

>>>>

@@

\section{Related Work}

Adding tolls to a Network Game has began as a standard practice to regulate the network flow, as it is well known that the equilibrium that occurs by the game is not guarranteed to be the most efficient.

TODO

@@

In this section we will describe the model utilised throughout this thesis.
If we attempted to call that model by including all of it's properties, the result would be a Heterogeneous Non-atomic 2-link Parallel Network Toll Congestion Pricing Game, so while we will describe each of those properties, we will eventually focus on heterogeneity and the pricing competition, thus calling it a Heterogeneous Pricing Game.

\section*{Non-atomic Parallel Network Games}

We consider a directed graph $G = (\{s, t\}, N)$ where $N = \{1,\dots, n\}$ is a set of parallel links from a source node $s$ to a target node $t$.
For the non-atomic case, there is one unit of traffic that wishes to travel from $s$ to $t$, described as the unit interval $[0, 1]$ endowed with Lebesque measure $\lambda$.
Each point $p \in [0, 1]$ will be called a \textit{player} and will be considered to be non-cooperative and contribute to the traffic by an infinitesimal amount.
As such the decisions of individual players have no effect on the game and it becomes natural to only consider non-zero measure collection of players.

A \textit{flow} is a (stochastic) vector $x = (x_i)_{i \in N}$ where $x_i$ is the total flow on link $i$.
The flow then creates congestion on the links, each described by a non-decreasing latency function $(\ell_i)_{i \in N}$ which we assume to be affine.
We denote by $\mathcal{L}_d$ the class of polynomial latency functions with nonnegative coefficients and degree at most $d$, and as such $\mathcal{L}_1$ becomes the class of affine latency functions.
Finally, the network can also be extended by allowing a set of tolls $t = (t_i)_{i \in N}$ to be assigned to each link, which in turn adds to the effective cost of a player using them.

The above set-up creates a Non-atomic Parallel Network Game with tolls where each player will select the edge that minimises their \textit{individual} traffic latency.
We consider two cases where players either react to tolls in a \textit{homogeneous} or a \textit{heterogeneous} manner.

\subsection*{Homogeneous players}

In the homogeneous case all players have an equal reaction to tolls, and therefore the effective cost of a player using link $i$ becomes $\ell_i(x_i) + t_i$.
For a given set of tolls $t$, a flow $x$ is a \textit{Wardrop equilibrium for $t$} if $\forall i, j \in N$ with $x_i > 0$ it holds that
\begin{equation}
	\label{wardrop_inequality_cost}
	\ell_i(x_i) + t_i \leq \ell_j(x_j) + t_j
\end{equation}
In that case, all links with $x_i > 0$ have equal effective costs, i.e. there exists some $K > 0$ such that $\ell_i(x_1) + t_i = K$ for all links $i$ with $x_i > 0$.
From a well-known result from Beckman et al. \cite{beckmann1956studies} and Dafermos and Sparrow \cite{1363388843888284416}, such and an equilibrium for $t$ exists and is unique, therefore for a toll vector $t$ we can denote by $x(t)$ the flow occurring on  that unique equilibrium for the given toll $t$.
We copy Lemma 2.1 from Harkes et al \cite{Harks_2019} as is for later use.
\begin{lemma}
	\label{lemma:wardrop_equilibrium}
	A flow $x$ is a Wardrop equilibrium for $t$ if and only if for all feasible flows $x^\prime$,
	\[\sum_{i \in N} (\ell_i(x_i) + t_i) \cdot (x_i - x_i^\prime) \leq 0\]
\end{lemma}
If $t = 0$ then the equilibrium is called the \textit{Wardrop equilibrium}.
It's also worth noting that looking at \eqref{wardrop_inequality_cost}, we can see that what actually matters, with regard to the flow, is the relative difference among the tolls.
I.e. if we consider any $c \in \R_+$ (or even $c \geq -\min_{i \in N} t_i$) then for $t^\prime = t + c$ the inequalities remain true for all $i, j \in N$ and thus the network game for the flow between tolls $t$ and $t^\prime$ is the same.

Finally for affine latencies we can calculate $x(t)$.
For $l_i(x_i) = a_i \cdot x_i + b_i, a_i > 0, b_i \geq 0$ and $t \in \R_+^N$ define $N(t) = \{i \in N | x_i(t) > 0\}$.
Since $x(t)$ is an equilibrium for $t$ and $\sum_{i \in N(t)}x_i(t) = 1$ we solve the equations with regard to the common efficient cost $K$ and then we get for all $i \in N(t)$
\begin{equation}
	x_i(t) = \frac{1 + \sum_{j \in N(t)}\frac{b_j + t_j - b_i - t_i}{a_j}}{\sum_{j \in N(t)}\frac{a_i}{a_j}}
\end{equation}

\subsection*{Heterogeneous players}

In the heterogeneous case, each player $p$ reacts differently to tolls, presumably due to different money/time valuation ratios.
We describe this by adding a money-sensitivity weight $\alpha(p)$ on the toll costs experienced by the player, making them see the cost of each edge as $c_i^p(x, t) = \ell_i(x_i) + \alpha(p) t_i$ and then seek the shortest link among them.
By also assuming that the players are sorted by money-sensitivity, we can describe their heterogeneity by defining a non-decreasing function $\alpha: [0, 1] \rightarrow [0, \infty]$.
We call $\alpha$ a \textit{distribution function}.
Even though the definition does allow functions that are not upper bounded with $\alpha(1) = +\infty$, we will always assume that $\alpha$ is finite on $[0, 1)$.
Also we will assume that $\alpha(0) > 0$, since, as we'll see in more detail in Section ??, allowing many players close to $0$ breaks the pricing game.

\section*{Pricing Games}

We introduce $n$ link owners which can assign a toll $t_i$ on link $i$ and profit from it.

\section*{2-link Pricing Games (our model)}
